---
title: "Coding_challenge_6"
author: "Bibek"
date: "2025-03-26"
output: 
  pdf_document:
            keep_md: yes
---
**#Question 1**
1) Regarding reproducibility, what is the main point of writing your own functions and iterations? 
Ans >> 

Writing own fucntions and iteration makes the code more organized and easier to run it again. With the function, we can code a set of instruction that can be used over again and again without rewriting them. Additionally, when we combine this with iteration which are loop structures it ensures that each steps in the analysis is clear and consistent for all the inputs. So, both of this ensure that our code behaves in same way, making it simpler to understand and free form minor error that might arise from copying and pasting. All of these ensures proper reproducibility.

**#Question 2**
2) In your own words, describe how to write a function and a for loop in R and how they work. Give me specifics like syntax, where to write code, and how the results are returned. 
Ans >>

a) For function

-First we give the name to function.Then we use "function ()" where we list variables that we want to pass i.e. inside the bracket.This is followed by the curly brackets where actual code is written for calculations or specific operation. Finally, we use "return()" to obtained specific output.
    
 Below is the example for this. When we call my_function(1), it returns 3 as output.
    
```{r}
##Example
my_function <- function(x) {
  result <- x + 2   ## This is where we write the code
  return(result)
}

my_function(1)
```
b) for loop

-The for loop starts with "for (i in sequence range) {}, where "i" is the iterator that uses the values from the sequence range one by one. The sequence is the list or range of values that the loop goes through one by one. It can be numbers, characters, or logical operators and guide the loop for how many times and in what order to run.The main code for execution is written inside the curly brackets. This code is executed by each value in the sequence which is specified by the "i". Then, print() function helps to obtained the result of the code in the console windows. This loop continues to run until the last values in the provided sequence. Instead of printing directly in console, we can use an empty objject before the loop and the store each value for later use.
  
  Below is the simple example for the for loop.(without null object)
  This will give the multiple of 2 till 20.
```{r}
for (i in 1:10){
  print(i*2)
}
```
  
**#Question 3**    
Read in the Cities.csv file from Canvas using a relative file path.
```{r}
Thecities <- read.csv("Sample_data/Cities.csv")
```

  
**#Question 4**  

Making the function
```{r}
distance <- function(lat1, lon1, lat2, lon2){
  rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  # Haversine formula
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a)) 

  # Earth's radius in kilometers
  earth_radius <- 6378137
  # Calculate the distance
  distance_km <- (earth_radius * c)/1000

return(distance_km)

}
```
**#Question 5** 
Computing the distance between Auburn, AL and New York City 
```{r}
# Get coordinates for Auburn
auburn <- subset(Thecities, city == "Auburn")
lat1 <- auburn$lat
lon1 <- auburn$long

# Get coordinates for New York City
nyc <- subset(Thecities, city == "New York")
lat2 <- nyc$lat
lon2 <- nyc$long

# Calculate the distance
distance_to_nyc <- distance(lat1, lon1, lat2, lon2)
distance_to_nyc
```
**#Question 6** 
Calculate the distance between all other cities in the data.
```{r}
#without creating null object:
for (i in seq_along(Thecities$city)){
    if (Thecities$city[i] != "Auburn") {
    lat2 <- Thecities$lat[i]
    lon2 <- Thecities$long[i]
    the_distance <- distance(lat1, lon1, lat2, lon2)
    print(the_distance)
    }
}

```
**Bonus Question:**
```{r}
#With null object to stored the data:

result <- NULL
#for loop
for (i in seq_along(Thecities$city)){          # Loop each row in the df 
    city_name <- Thecities$city[i]             # Extract city name
    if (city_name != "Auburn") {               # Skip Auburn itself
    lat2 <- Thecities$lat[i]                   # Extract the latitude and longitude
    lon2 <- Thecities$long[i]
    the_distance <- distance(lat1, lon1, lat2, lon2)  # Using above distance function to calculate distance from Auburn
    print(the_distance)                     # Print the distance in the console
    
    # making new row to results as mentioned in the question
    new_row <- data.frame(City1 = Thecities$city[i] , City2 = "Auburn", Distance_km = the_distance)
    result <- rbind(result, new_row)
  
  }
}

result
```
*Without using if function; this will not check if the same cities are being compared with each other. So we will get Auburn to Auburn at zero distance apart.*
```{r}

result_distance2 <- NULL 

for (i in seq_along(Thecities$city)) { 
      lat1 <- Thecities$lat[Thecities$city=="Auburn"]
      lon1 <- Thecities$long[Thecities$city=="Auburn"]
      lat2 <- Thecities$lat[i]
      lon2 <- Thecities$long[i]

      the_distance <- distance(lat1, lon1, lat2, lon2)  # Using above distance function to calculate distance from Auburn
    print(the_distance)                     # Print the distance in the console
    
    # making new row to results as mentioned in the question
    new_row <- data.frame(City1 = Thecities$city[i], City2 = "Auburn", Distance_km = the_distance)
    result_distance2 <- rbind(result_distance2, new_row)
    
}
result_distance2
```


**#Question 7** 
Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 6. 

Please click this:
[**Link to my GitHub**](https://github.com/BZD0052/Class_example/tree/main/Coding_Assignment/Function_and_iteration)








