---
title: "Function and Iteration Homework"
author: "Bibek"
date: "2025-03-26"
output: 
  pdf_document:
            keep_md: yes
---
Iteration and functions
```{r}
## Required packages
library(ggplot2)
library(drc) #dose response curve
library(tidyverse)
```
Function:
Reusable block of code designed to perform a specific task.
```{r}
#Without function #(5*(degree_f - 32)/9)
(5*(32 - 32)/9)
(5*(39 - 32)/9)
(5*(40 - 32)/9)

#Converting Fahrenheit to Celcius
F_to_C <- function(f_temp){
  celsius <- (5*(f_temp - 32)/9)
  return(celsius)
}
```
Function structure model:

sample.function.name <- function(.... variable.....){
  .... main code here.... 
   return(... output ...)
}

Example of function
```{r}
#Using the fucntion
F_to_C(32)

Temp <- c(32, 39, 40, 44, 42, 53, 63)
celcius_scale <- F_to_C(Temp)

#Result 
celcius_scale
```
Iteration:
Repeateadly using a block of code such as a loop.
Helps reduce copying and pasting errors.
```{r}
#rep : this function allows you to repeat elements easily
rep("A", 3)
rep(c("A", "B"), 20)
rep(c("Bibek", "Rocks"), 5)

rep(c(1, 2, 3), 3, each =3)

#seq : this helps to write sequences of numbers easily
1:7
seq(from=1, to =7)
seq(from=0, to =20)
seq(from=1, to =30, by=2)

# combine rep and seq

rep(seq(from=0, to =5, by=2), times=3, each=2)

#seq_along : this allows to generate a sequence of numbers based on non-integer values. 

LETTERS
seq_along(LETTERS)

# THE for loop

for (i in 1:10){
  print(i*2)
}

#more complicated
for (i in -10:100){
  result <- F_to_C(i)
  print(result)
}

#to save the value 

celcius.df <- NULL # creat a null object
for (i in -10:100){   
  result <- data.frame(F_to_C(i), i)   #save the result in data frame at each iteration 
  celcius.df <- rbind.data.frame(celcius.df, result) #row bind with celcius.df
}

celcius.df

```
Practical example fungicide sensitivity data.
```{r}
EC50.data <- read.csv("Sample_data/EC50_all.csv")

isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == "ILSO_5-41c"] ~ 
                  EC50.data$conc[EC50.data$is == "ILSO_5-41c"], 
                fct = LL.4(fixed = c(NA, NA, NA, NA), 
                           names = c("Slope", "Lower", "Upper", "EC50")), 
                na.action = na.omit)
# outputs the summary of the paramters including the estimate, standard
# error, t-value, and p-value outputs it into a data frame called
# summary.mef.fit for 'summary of fit'
summary.fit <- data.frame(summary(isolate1)[[3]])
# outputs the summary of just the EC50 data including the estimate, standard
# error, upper and lower bounds of the 95% confidence intervals around the
# EC50
EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
           interval = "delta")[[1]]


##loop for above code

nm <- unique(EC50.data$is)
nm


for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
                    EC50.data$conc[EC50.data$is == nm[[i]]], 
                  fct = LL.4(fixed = c(NA, NA, NA, NA), 
                             names = c("Slope", "Lower", "Upper", "EC50")), 
                  na.action = na.omit)
  print(nm[[i]])
  summary.fit <- data.frame(summary(isolate1)[[3]])
  EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
             interval = "delta")[[1]]
  EC50
}

```
Saving each iteration output for above code
```{r}
EC50.ll4 <- NULL # create a null object 
for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
                    EC50.data$conc[EC50.data$is == nm[[i]]], 
                  fct = LL.4(fixed = c(NA, NA, NA, NA), 
                             names = c("Slope", "Lower", "Upper", "EC50")), 
                  na.action = na.omit)
  print(nm[[i]])
  summary.fit <- data.frame(summary(isolate1)[[3]])
   EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
             interval = "delta")[[1]]
  EC50
  isolate.ec_i <- data.frame(nm[[i]], EC50) # create a one row dataframe containing just the isolate name and the EC50
  colnames(isolate.ec_i) <- c("Isolate", "EC50") # change the column names
  # Then we need to append our one row dataframe to our null dataframe we created before
  # and save it as EC50.ll4. 
  EC50.ll4 <- rbind.data.frame(EC50.ll4, isolate.ec_i)
}
EC50.ll4
```

Using tidyverse
```{r}
EC50.data %>%
  group_by(is) %>%
  nest() %>%  # allow for sub dataframe within a dataframe
  mutate(ll.4.mod = map(data, ~drm(.$relgrowth ~ .$conc,  # create a ll4 model column equal to summary output of drm function over data
                                   fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                              names = c("Slope", "Lower", "Upper", "EC50"))))) %>%
  mutate(ec50 = map(ll.4.mod, ~ED(., 
                                  respLev = c(50), 
                                  type = "relative",
                                  interval = "delta")[[1]])) %>%
  unnest(ec50) ##see the output for ec50
```

