---
title: "Coding challenge 5-Data Wrangling"
author: "Bibek"
date: "2025-03-20"
output: 
  pdf_document:
            keep_md: yes
            toc: TRUE    # add table of content

---
**#Question 1**

#Loading dataset
```{r}

Diversity <- read.csv("Sample_data/DiversityData.csv")
Meta <- read.csv("Sample_data/Metadata.csv")
```

**#Question 2 **
#Loading library
#Join them together via common column 'code'.
```{r}
library(tidyverse)
alpha <- left_join(Diversity, Meta, by = "Code")
```


**#Question 3**
# Calculate Pielou's evennes index. This equals to the Shannon diversity index (column Shannon) divided by the log of the richness column. 
```{r}
aplha.even <- mutate(alpha, Pielou_Eveness_Index= shannon/log(richness)) # creating a new column Pielou_Eveness_Index
```

**#Question 4**
#summarise function for Mean and SE of grouped by crop over time.
```{r}
alpha_average <- aplha.even %>%
  group_by(Crop, Time_Point) %>% # grouping by crop and time point
  summarise(Mean.even = mean(Pielou_Eveness_Index), # calculating the mean even, count, stdeviation, and standard error
            count = n(), 
            sd.even = sd(Pielou_Eveness_Index),
            std.err = sd.even/sqrt(count))
alpha_average
```
**#Question 5**
# Calulate the differences between soybean column, soil column, and the difference between the cotton column and the soil column.
```{r}
alpha_average2 <- alpha_average %>%
  select(Time_Point, Crop, Mean.even) %>%
  pivot_wider(names_from = Crop, values_from = Mean.even) %>% # pivot to wide format
  mutate(diff.cotton.even = Soil - Cotton,
         diff.soybean.even = Soil - Soybean) ##  calculate the difference between the mean between the groups.
```

**#Question 6**
# Connecting plot

```{r Difference from soil in Pielou eveness vs Time}
alpha_average2 %>%
  select(Time_Point,diff.cotton.even,diff.soybean.even)%>%
  pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff") %>%
  ggplot(aes(x=Time_Point, y= value, color = diff)) +
  geom_line()+
  theme_classic()+
  xlab("Time (hrs)") +
  ylab("Difference from soil in Pielou's eveness")

```
**#Question 7**
#Make .md file, push to GitHub isnide directory called Coding challenge 5 and provide link in you submission file.
#See Coding challenge 5 folder.

Peasle click this:
[**Link to my GitHub**](https://github.com/BZD0052/Class_example.git)
