---
title: "Coding challenge 4 - R_markdown"
author: "Bibek"
date: "2025-02-27"
output:
  word_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
  md_document:
    variant: gfm
    toc: true
  Word_document:
    toc: true
  pdf_document:
    toc: true
---

#  Link

**Link to the manuscript where the data are published**

[**Noel, Z.A., Roze, L.V., Breunig, M., Trail, F. 2022. Endophytic fungi as promising biocontrol agent to protect wheat from Fusarium graminearum head blight. Plant Disease. https://doi.org/10.1094/PDIS-06-21-1253-RE **](https://doi.org/10.1094/PDIS-06-21-1253-RE)



**Questions 1**


a.	YAML header

In R markdown, YAML header is a section at the top of the document that contains metadata about the file. It is enclosed between "---" lines and helps to tell how to format the output file.

For example, in this markdown i have YAML hear  with infromation like title, author, date which will appearin the rendered files. I have also specified output format to generate markdown file using gfm. Other than this i have also other output options which i can choose while knitting.

b.	Literate programming

Literate programming means writing both the code and explanation in the same document whihch make analysis easy to understand and reproduce.




**Question 2**


Here i will use the relative path of the Mycotoxin.csv data file with na.strings option set to "na".
I will be using the codes from the coding challenge 3 assignment.

```{r Reading CSV}
#Read .csv
#using relative path
DON.database <- read.csv("Sample_data/MycotoxinData.csv", na.strings = "na")
str(DON.database) #visualize the structure of the data frame.
```
Now, I will separate code chunk for the figures plotting the DON data, 15ADON, and Seedmass, and one for the three combined using ggarrange. 

# First level header

This is for DON data.

```{r Plotting DON}
library(ggplot2) # loading the ggplot2 package
library(ggpubr) ##integrating multiple statistics and plots

cbbPalette <-c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
myCols <- cbbPalette[c(3, 4)]  ## choosing 3rd and 4th color palette from the cbbPalette.

##Question 1####
#Create a boxplot for DON (deoxynivalenol) values
DON <- ggplot(DON.database, aes(x = Treatment, y = DON, fill = Cultivar)) +
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA ) +   # Create boxplots with dodge positioning; remove default outlier points.
  geom_point(position = position_jitterdodge(0.3), shape = 21, color ="black", alpha =0.6) +  #jittered points to show individual observations
  scale_fill_manual(values = myCols) + ## Use custom fill colors for the Cultivar groups.
  ylab("DON(ppm)") + # Label the y-axis and remove x-axis label.
  xlab("") +
  theme_classic() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Cultivar) # Create separate panels for each Cultivar.

DON #Visual

#Reorder the factor order level
DON.database$Treatment <- factor(DON.database$Treatment, levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70"))

# Re-displaying the plot will now use the new Treatment order.
DON <- ggplot(DON.database, aes(x = Treatment, y = DON, fill = Cultivar)) +
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA ) +   # Create boxplots with dodge positioning; remove default outlier points.
  geom_point(position = position_jitterdodge(0.3), shape = 21, color ="black", alpha =0.6) +  #jittered points to show individual observations
  scale_fill_manual(values = myCols) + ## Use custom fill colors for the Cultivar groups.
  ylab("DON(ppm)") + # Label the y-axis and remove x-axis label.
  xlab("") +
  theme_classic() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Cultivar) # Create separate panels for each Cultivar.

```

# Second level header

This id for 15ADON plot.

```{r Plotting 15ADON}
#Change the y-variable to plot X15ADON and MassperSeed_mg. 

# Boxplot for X15ADON values
X15ADON <- ggplot(DON.database, aes(x = Treatment, y = X15ADON, fill = Cultivar)) +
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) +
  geom_point(shape = 21, color = "black", position = position_jitterdodge(0.3), alpha = 0.6) +
  scale_fill_manual(values = myCols) +
  ylab("15ADON") + # y-axis label for X15ADON
  xlab("") +
  theme_classic() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Cultivar)
  
X15ADON # Display the X15ADON plot
```



# Third level header

This id for Seedmass plot.
```{r Plotting MassperSeed_mg}
# Boxplot for MassperSeed_mg (Seed Mass in mg)

MassperSeed_mg <- ggplot(DON.database, aes(x = Treatment, y = MassperSeed_mg, fill = Cultivar)) +
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) +
  geom_point(shape = 21, color = "black", position = position_jitterdodge(0.3), alpha = 0.6) +
  scale_fill_manual(values = myCols) +
  ylab("Seed Mass (mg)") +  # y-axis label for Seed Mass
  xlab("") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Cultivar)

MassperSeed_mg #Display the MassperSeed_mg plot
```

# Fourth level header

This is combined 3 plots using ggarrange.


```{r Combined plot}
# Combining all plots together

Fig_combo<-ggarrange(    ## Arrange the three plots (DON, X15ADON, MassperSeed_mg) side by side in one row
  DON,
  X15ADON,
  MassperSeed_mg,
  labels = "auto", # Add labels (a, b, c) to each plot
  nrow = 1,  # Arrange in one row
  ncol = 3,   # Three columns
  legend = FALSE,   # Do not display the legend in the combined plot
  common.legend = T    #If TRUE, a common unique legend will be created for arranged plots.
)

Fig_combo
```















