---
title: "Coding_challenge_7_linear_model"
author: "Bibek"
date: "2025-04-03"
output: 
  html_document:
    keep_md: yes
    toc: true
    toc_float: true
  pdf_document:
            keep_md: yes
---

# Question 1

**Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor**
 *Example shown here.*
 *STAND$Treatment <- as.factor(STAND$Treatment)*

```{r Question 1}
#Loading libraries
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)


Emergence<-read.csv("Sample_data/PlantEmergence.csv", na.strings="na") #Loading the plant emergence data 

#Converting treatment, days after lanting and replication into factors
Emergence$Treatment <- as.factor(Emergence$Treatment) 
Emergence$DaysAfterPlanting<-as.factor(Emergence$DaysAfterPlanting)
Emergence$Rep<-as.factor(Emergence$Rep)
```

# Question 2

**Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results*. **

```{r Q3:Linear model with interaction}
lm1<-lm(Emergence~Treatment*DaysAfterPlanting, data=Emergence)
summary(lm1)
anova(lm1)
```


# Question 3

**Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.**

*ANs:*

Based on the results of the linear model in question 2,the interaction is not significant. So, we don't need to fit the interaction.

```{r}
lm2<-lm(Emergence~Treatment+DaysAfterPlanting,data=Emergence)
summary(lm2)
anova(lm2)
confint(lm2)
```

We found that Treatment 2 had 134.53 (+-6.775364;+-95% CI)  units less emergence time than Treatment 1 (p-value <2e-16 ***).
The intercept represents the average emergence for treatment 1 at 7 days after planting which is 182.163 days.
The coefficient of Treatment 2 is the difference in emergence between treatment 2 and treatment 1.

# Question 4
**Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.**

```{r emmeams}
lsmeans<-emmeans(lm2,~Treatment)
results_lsmeans<-cld(lsmeans,alpha=0.05,details=TRUE) #compact letter display
results_lsmeans
```

Based on the estimated marginal means adjusted for Days After Planting, Treatment 2 had the lowest emergence (55.6) and was significantly different from all other treatments (p < 0.0001). Treatments 3, 5, and 6 showed the highest average emergence (~200) and were statistically similar, forming the top-performing group.Treatments 1, 4, 8, and 9 exhibited moderate emergence (~190–198) and were not significantly different from each other, grouping together. Treatment 7 had a moderate emergence level (188.0) but statistically similar to treatments 1,4,8, and 9. In conclusion, Treatment 2 greatly reduced emergence, while the remaining treatments performed similarly, with Treatments 3, 5, and 6 showing the highest emergence.

# Question 5

**The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters.**

```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}


plot_cldbars_onefactor(lm2,"Treatment")

```

The letters above the bars (like "a", "ab", "b", "c") come from a Tukey’s Honest Significant Difference (HSD) test (α = 0.05). This test compares the mean number of emerged plants across all treatments and tells you which treatments are statistically different from each other. Treatments that share the same letter are not significantly different from one another in terms of the number of emerged plants. For example, Treatments labeled with "a" or "ab" performed similarly, while those with different letters (like "a" vs. "b" or "c") are significantly different. Treatment 2 (labelled with"c") , had the lowest emergence and was significantly different from all others. Treatments 3, 5, and 6 (labeled with "a"), had the highest emergence. Treatments 1, 4, 8, and 9 (labelled with "ab") fall in between and are not statistically different from either the highest or middle group. These groupings help quickly identify which treatments had similar or significantly different effects on plant emergence.

# Question 6

**Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the.md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.**

#See Coding challenge 6 folder.

Peasle click this:
[**Link to my GitHub**](https://github.com/BZD0052/Class_example/Coding_Assignment/Linear_model)

