---
title: "Data Wrangling Homework"
author: "Bibek"
date: "2025-03-18"
output: 
      pdf_document:
        keep_md: yes

---
Tidyverse is a bunch of packages and fucntion for easiers use case of large datset in R.
This assignments will cover follwing headings:
-Data wrangling & manipulation
-mutate()
-select()
-filter()
-the pipe %>%
-summarise()
-group_by()
-joining
-pivoting
-Integration with plotting
               
###Loading the dataset
```{r}
microbiome.fungi <- read.csv("Sample_data/Bull_richness.csv")
str(microbiome.fungi)

library(tidyverse)
```
#####Select() function- choose certain columns of th data.
```{r}
microbiome.fungi2 <- select(microbiome.fungi, SampleID, Crop, Compartment:Fungicide, richness)
```

##### filter()
Function for sub-setting data easily.

```{r}
#Simple filtering
head(filter(microbiome.fungi2, Treatment == "Conv."))

#Complex filtering (&)
head(filter(microbiome.fungi2, Treatment == "Conv." & Fungicide == "C"))

# Another more complex (|)
head(filter(microbiome.fungi2, Sample == "A" | Sample == "B")) # samples A or B

```

#### mutate()
Helps to create new columns quickly.
```{r}
# Create a new column called logRich
head(mutate(microbiome.fungi2, logRich = log(richness)))

#Previous way to do it
#microbiome.fungi2$logRich <- log(microbiome.fungi2$richness)

# Creating a new column with combined Crop and Treatment 
head(mutate(microbiome.fungi2, Crop_Treatment = paste(Crop, Treatment)))
```

#### The pipe, %>% 
Helps to combine the various function together.

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # sub-setting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column log richness
  head() # displaying
```
#### summarise()
Helps to find out means and standard deviation or errors.

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # sub-setting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column log richness
  summarise(Mean.rich = mean(logRich)) # calculate overall mean log richness within the conventionally managed treatment

# Can also connect multiple summary statistics
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # sub-setting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column log richness
  summarise(Mean.rich = mean(logRich),
            n= n(),  # calsulates the number of n
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
  


```

#### group_by()

For obtaining the stat by group.
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% 
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>%
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
```
####adding ggplot to the previous using pipe

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% 
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>%
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  ggplot(aes(x = Fungicide, y = Mean.rich)) + # adding in a ggplot
  geom_bar(stat="identity") +
  geom_errorbar( aes(x=Fungicide, ymin=Mean.rich-std.err, ymax=Mean.rich+std.err), width=0.4) +
  theme_minimal() +
  xlab("") +
  ylab("Log Richness") +
  facet_wrap(~Treatment)
```
#### Joining 

Allow to combine multiple data set based on common set of variables.

They include:
 - left_join()
  - Keep all rows of X and add matching rows from Y. Any rows in Y that don’t match X are excluded.
 - right_join() 
  - reverse of left_join() 
 - inner_join() 
  - only keep rows that are common to both X AND Y, remove everything else. 
 - full_join() 
  - Keep any columns that are in either X or Y

```{r}
# Examples
# selecting just the richness and sample ID
richness <- microbiome.fungi %>%
  select(SampleID, richness)

# selecting columns that don't include the richness
metadata <- microbiome.fungi %>% 
  select(SampleID, Fungicide, Crop, Compartment, GrowthStage, Treatment, Rep, Sample)

head(richness)
head(metadata)


# using leftjoin and adding the richness data to the metadata based on on the common column of sampleID
head(left_join(metadata, richness, by = "SampleID"))
```
#### Pivoting
Used for converting from wide to long format and back again. we can do this using 'pivot_longer()' and 'pivot_wider()'.
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting column
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide 
  summarise(Mean = mean(richness)) %>% # calculate the mean
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
## can now easily calculate the difference between the mean between the fungicide and control groups.
  mutate(diff.fungicide = C - F)
  
```
#### adding a plot to above chunk
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting column
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide 
  summarise(Mean = mean(richness)) %>% # calculate the mean
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
## can now easily calculate the difference between the mean between the fungicide and control groups.
  mutate(diff.fungicide = C - F) %>%
  ggplot(aes(x= Treatment, y = diff.fungicide)) + 
  geom_col() +
  theme_minimal() +
  xlab("") +
  ylab("Difference in average species richness")
```








